Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-04 Master Thesis/synopsis.tex
Add proof for 2.4 for non-finitely-generated fields

@@

\begin{proof}
The degree $[K(\mu_\ell, \sqrt{-1}):K]$ is prime to $\ell$, thus we can assume that $\mu_\ell\subset K$ and $K$ is not formally real with a usual norm argument. Now we can replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2) = \mu_\ell^{\otimes 2}$.

We write $\mathrm{H}^3(L) = \mathrm{H}^3(L, \mathbb{Z}/\ell\mathbb{Z}(2))$. Let $\mathfrak{X}$ be a model of $K$. $\mathfrak{X}_i$ denotes the set of points $x\in\mathfrak{X}$ whose closure $\overline{\{x\}}$ is $i$-dimensional. For $x\in\mathfrak{X}_1$ the local ring $\mathcal{O}_{\mathfrak{X}, x}$ is a discrete valuation ring and defines a valuation $v_x\in\operatorname{Spv}_\text{d}^{(1)}(K)$. By [K1] Theorem 0.8, the following map
\[ \mathrm{H}^3(K)\longrightarrow \prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_{v_x}) \]
is injective and its image is contained in $\bigoplus \mathrm{H}^3(K_{v_x})$. By [K1] Lemma 1.4, for a discrete valuation $v$ of $K$ the map
\[ \mathrm{H}^3(K_v) \longrightarrow \prod_w\mathrm{H}^3(K_w) \]
where $w$ ranges over all rank-$2$ valuations with $v(w) = v$, can be identified with the following map:
\[ {}_\ell\operatorname{Br}(k) \longrightarrow \prod_{\tilde{w}\in\operatorname{Spv}^{(1)}(k)} {}_\ell\operatorname{Br}(k_{\tilde{w}}) \]
where $k = k(v)$. This map is injective by the classical Hasse principle and its image is contained in $\bigoplus {}_\ell\operatorname{Br}(k_{\tilde{w}})$. Combining these maps yields the injection:
\[ \mathrm{H}^3(K)\longrightarrow \bigoplus_{x\in\mathfrak{X}_1}\  \bigoplus_{\substack{w\in\operatorname{Spv}_\text{rd}^{(2)}(K) \\ v(w) = v_x}} \mathrm{H}^3(K_w) \]
The direct sum ranges over a subset of $|\operatorname{Spv}(K)|$, thus this concludes the proof.
\end{proof}

>>>>

\begin{lemma}\label{2.4-pre}
Let $K$ be finitely generated, $\ell\neq 2$ be a prime and $f\in\mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z})$. Then the support of $f$ is closed in $|\operatorname{Spv}(K)|$:
\[ \operatorname{supp}(f) = \{w\in |\operatorname{Spv}(K)|\mid \operatorname{res}_{K_w/K}(f)\neq 0 \} \]
\end{lemma}

@@

\begin{proof}
The degree $[K(\mu_\ell, \sqrt{-1}):K]$ is prime to $\ell$, thus we can assume that $\mu_\ell\subset K$ and $K$ is not formally real with a usual norm argument. Now we can replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2) = \mu_\ell^{\otimes 2}$. We write $\mathrm{H}^3(L) = \mathrm{H}^3(L, \mathbb{Z}/\ell\mathbb{Z}(2))$.

First, let's consider the case where $K$ is finitely generated. Let $\mathfrak{X}$ be a model of $K$. $\mathfrak{X}_i$ denotes the set of points $x\in\mathfrak{X}$ whose closure $\overline{\{x\}}$ is $i$-dimensional. For $x\in\mathfrak{X}_1$ the local ring $\mathcal{O}_{\mathfrak{X}, x}$ is a discrete valuation ring and defines a valuation $v_x\in\operatorname{Spv}_\text{d}^{(1)}(K)$. By [K1] Theorem 0.8, the following map
\[ \mathrm{H}^3(K)\longrightarrow \prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_{v_x}) \]
is injective and its image is contained in $\bigoplus \mathrm{H}^3(K_{v_x})$. By [K1] Lemma 1.4, for a discrete valuation $v$ of $K$ the map
\[ \mathrm{H}^3(K_v) \longrightarrow \prod_w\mathrm{H}^3(K_w) \]
where $w$ ranges over all rank-$2$ valuations with $v(w) = v$, can be identified with the following map:
\[ {}_\ell\operatorname{Br}(k) \longrightarrow \prod_{\tilde{w}\in\operatorname{Spv}^{(1)}(k)} {}_\ell\operatorname{Br}(k_{\tilde{w}}) \]
where $k = k(v)$. This map is injective by the classical Hasse principle and its image is contained in $\bigoplus {}_\ell\operatorname{Br}(k_{\tilde{w}})$. Combining these maps yields the injection:
\[ \mathrm{H}^3(K)\longrightarrow \bigoplus_{x\in\mathfrak{X}_1}\  \bigoplus_{\substack{w\in\operatorname{Spv}_\text{rd}^{(2)}(K) \\ v(w) = v_x}} \mathrm{H}^3(K_w) \]
The direct sum ranges over a subset of $|\operatorname{Spv}(K)|$, thus this shows the assertion. Now, we drop the assumption that $K$ is finitely generated.

Let $f\in\mathrm{H}^3(K)$ such that $\operatorname{res}_{K_w/K}(f) = 0$ for all $w\in |\operatorname{Spv}(K)|$. There exists a finitely generated subfield $K_0$ of $K$ such that $f = \operatorname{res}_{K/K_0}(g)$ for some $g\in\mathrm{H}^3(K_0)$ and $K/K_0$ is algebraic. Let $\{K_i\mid i\in I\}$ be the set of all finite subextensions of $K/K_0$. For $w\in |\operatorname{Spv}(K)|$ we denote the restriction of $w$ to $K_i$ with $w$ as well.

Let  $w\in |\operatorname{Spv}(K)|$. Since $\mathrm{H}^3(K_w) = \varinjlim_I \mathrm{H}^3((K_i)_w)$ there exists an $i = i(w)\in I$ such that $\operatorname{res}_{(K_i)_w/K}(g) = 0$. By Lemma \ref{2.4-pre} there is an open neighborhood $U_{i(w)}\subset |\operatorname{Spv}(K_i)|$ of $w$ such that $\operatorname{res}_{(K_i)_v/K}(g)=0$ for all $v\in U_{i(w)}$. For $K_i\subset K_j\subset K$ let
\[ \varphi_i: |\operatorname{Spv}(K)| \to |\operatorname{Spv}(K_i)| \]
\[ \varphi_{ij}: |\operatorname{Spv}(K_j)|\to |\operatorname{Spv}(K_i)| \]
be the restriction maps. They are continuous. Since $|\operatorname{Spv}(K)|$ is quasicompact we can find a finite set $S\subset |\operatorname{Spv}(K)|$ such that:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)}(U_{i(w)}) = |\operatorname{Spv}(K)| \]
Let $j\in I$ such that $K_j$ is the composite of the fields $K_{i(w)},\ w\in S$. We also have:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)j}(U_{i(w)}) = |\operatorname{Spv}(K_j)| \]
Let $v\in |\operatorname{Spv}(K_j)|$ and let $i = i(w),\ w\in S$ such that $v\in\varphi_{ij}^{-1}(U_{i(w)})$. We have:
\[ \operatorname{res}_{(K_j)_v/K_j}(\operatorname{res}_{K_j/K_0}(g)) = \operatorname{res}_{(K_j)_v/(K_i)_v}(\operatorname{res}_{(K_i)_v/K_0}(g)) = 0 \]
Since $K_j$ is finitely generated and we proved the statement for finitely generated fields, we obtain $\operatorname{res}_{K_j/K_0}(g) = 0$, thus $f = \operatorname{res}_{K/K_0}(g) = 0$.
\end{proof}
